---
title: "Protest and Government Crisis"
author: "Conor Cunningham"
description: "Correlation between civic unrest and government crisis"
output:
  distill::distill_article:
    self_contained: false
---





##Does the amount of anti government protests correlate with the number of government crisis? From first glance it would seem that this would be the logical assumtion. I want to determine if this is really the case. I will use the data set "Political Violence" to determine this. My two variables will be "govtcrisis =  Major Government Crises: Any rapidly developing situation" and "agovdem = antigovernment demonstrations." If it is indeed the case, I would like to know to what extent they are corelated. How much does the quantity of anti-government demonstrations corelate with the quantity of government crisisis. I hypothesis that the amount of anti-government crisis WILL correlate with an increase in the amount anti-government protests. Therefore this will be a positive regression with the number of anti-government protests being the dependent variable and the number of government crisis being the independent variable. To prove this I will look to see if there is a greater number of government crisis when there is an increase in the amount of anti-government protests. The data set is made up of 208 identified government structures indicated by "country." For each variable there are also three data sets divided by year and I will look at all three (60's,70's,80,s) together. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse)
```

```{r, fig.width = 11, fig.height = 4}
plot_ <- violence%>%
  ggplot(aes(x=democ70,y=riots70))+
  geom_point()+
  geom_smooth()+
  xlab("Measure of Democracy in the 70s")+
  ylab("Number of Riots in the 70s")

plot_
```
govtcris60

```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=riots60,y=govtcris60))+
  geom_point(position = "jitter")+
  geom_smooth()+
  xlab("Number of riots in the 60s")+
  ylab("Number of Government Crisis in the 60s")

plot_
```


```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=riots60,y=surp60))+
  geom_point(position = "jitter")+
  geom_smooth()+
  xlab("Number of riots in the 60s")+
  ylab("Fiscal Surplus/GDP: Decade average of ratio of central government surplus")

plot_
```

```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=riots60,y=newspaperPC60))+
  geom_point(position = "jitter")+
  geom_smooth()+
  xlab("Number of riots in the 60s")+
  ylab("Newspapers per capita")

plot_
```

```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=riots60,y=tvPC60))+
  geom_point(position = "jitter")+
  geom_smooth()+
  xlab("Number of riots in the 60s")+
  ylab("TVs per capita in the 60s")

plot_
```

```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=riots70,y=democ70))+
  geom_point(position = "jitter")+
  geom_smooth()+
  xlab("Number of riots in the 70s")+
  ylab("Quality of democracy in the 70s")

plot_
```


```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=tvPC70,y=democ70))+
  geom_point(position = "jitter")+
  geom_smooth()+
  xlab("TVs per capita in the 70s")+
  ylab("Quality of the democracy in the 70s")

plot_
```


```{r, fig.width = 8, fig.height = 3}
plot_ <- violence%>%
  ggplot(aes(x=newspaperPC60,y=riots60))+
  geom_point()+
  geom_smooth()+
  xlab("Number of riots in the 60s")+
  ylab("Number of Government Crisis in the 60s")

plot_
```


```{r, fig.width = 11, fig.height = 4}
plot_ <- 
ggplot(data = violence, mapping = aes(x = agovdem60, y = govtcris60)) +
geom_line(mapping = aes(group = country), color = "gray70") +
geom_smooth(method = "loess", se = FALSE, size=1.1) +
labs (x = "antigovernment protests", y = "Major Governemnt Crisis", title = "Antigovernment and Governmental Crisis", caption = "Source: Violence") +
facet_wrap(~continent, nrow = 1) 


plot_ 
```

